---
title: "Example Code for Challenges"
author: "Coded by Allison"
date: "March 2, 2020"
output: html_document
---


# Challenge 2 
## Part 3 

What was the heaviest animal measured in each year? Return the columns `year`, 
`genus`, `species_id`, and `weight`.


```{r, answer=TRUE, purl=FALSE}
surveys %>%
  filter(!is.na(weight)) %>%
  ## I needed to remove the NA's 
  group_by(year) %>%
  filter(weight == max(weight)) %>%
  ## This comparison fails, if you have NA's in your variable! 
  select(year, genus, species, weight) %>%
  arrange(year)
```


# Challenge 4  
## Part 1  

> 1. Pivot the `combined` data frame to a wide format, with `year` as columns, 
>   `plot_id` as rows, and the number of genera per plot as the values. You will
>   need to summarize before reshaping, and use the function `n_distinct()` to 
>   get the number of unique genera within a particular chunk of data. It's a 
>   powerful function! See `?n_distinct` for more.

```{r, answer=TRUE, purl=FALSE}
surveys_wide_genera <- combined %>%
  group_by(plot_id, year) %>%
  summarize(n_genera = n_distinct(genus)) %>%
  pivot_wider(names_from = year, values_from = n_genera)
 
View(surveys_wide_genera)
```

## Part 2 

> 2. Now take that data frame and use `pivot_longer()` to pivot it to the long 
>    format it was in before, so that each row is a unique `plot_id` by `year` 
>    combination.


```{r, answer=TRUE, purl=FALSE}
surveys_wide_genera %>%
  pivot_longer(cols = -plot_id, names_to = "year", values_to = "n_genera")

## first argment is what columns you want to gather (- means everything but this column)
## second argment is what the name of the new column (with the pivoted columns) will be
## third argument is what the name of the new value column (with the values of the cells) will be
```


## Part 3 

> 3. The `complete` data set has two measurement columns: `hindfoot_length` and
>    `weight`.  This makes it difficult to do things like look at the 
>    relationship between mean values of each measurement per year in different
>    plot types. Let's walk through a common solution for this type of problem.
>    First, use `pivot_longer()` to create a dataset where we have a names 
>    column called `measurement` and a values column that takes on the value of 
>    either `hindfoot_length` or `weight`. 
>
>    *Hint*: You'll need to specify which columns to pivot into longer format!

```{r, answer=TRUE, purl=FALSE}
combined_long <- combined %>%
  pivot_longer(cols = hindfoot_length:weight, 
               names_to = "measurement", 
               values_to = "value")

## If the columns were not sequential we woud instead have used: 
## cols = c(hindfoot_length, weight)

## Remember: when you pivot columns to a long format, you have to create two 
## new columns!
## First: where the names of the pivoted columns go (names_to)
## Second: where the values inside the pivoted columns go (values_to)  
```


## Part 4 

> 4. With this new data set, calculate the average of each `measurement` in each
>    `year` for each different `plot_type`. Then pivot these summaries into a 
>    a data set with a column for `hindfoot_length` and `weight`. 
> 
>   *Hint*: This sounds like you want to pivot the data to be a wider format! 


```{r, answer=TRUE, purl=FALSE}
combined_long %>%
   group_by(year, measurement, plot_type) %>%
   ## group by three variables! 
   summarize(mean_value = mean(value, na.rm=TRUE)) %>%
   ## summarize the mean of each measurement at each plot type, each year 
   pivot_wider(names_from = measurement, values_from = mean_value)
   ## pivot the dataframe into a wider format, using the measurements as a pivot
   ## and filling each column with the mean_value for that measurement/plot type/year
```

